Joint estimation of phylogenies and ancestral states is not consistent

### 2018-01-30 joint_inf.tex
comments

An estimator is said to be statistically consistent if it converges to the generating model with probability 1 in the large-data limit; existing consistency proofs for maximum likelihood phylogenetics \cite{RoyChoudhury2015-ta} apply only to estimating model parameters when the ancestral sequences have been integrated out of the likelihood.
These proofs do not readily extend to include estimating ancestral states.
Moreover, examples of inconsistency arising from problems where the number of parameters increases with the amount of data \cite{Neyman1948-tt} indicate that joint inference of trees and ancestral states may not enjoy good statistical properties.
In this case those additional parameters come in the form of the states of ancestral sequences.
%EM I don't see this sentence in the most recent version of the paper via search. Can you point me to a section?
Although \cite{Sagulenko2017-jo} explicitly warn that their approximation is for the case where ``branch lengths are short and only a minority of sites change on a give branch,'' their work motivates understanding when exactly this approximate technique breaks down, even when sequence data is ``perfect,'' i.e., is generated without sampling error according to the exact model used for inference.

>>>>

An estimator is said to be statistically consistent if it converges to the generating model with probability 1 in the large-data limit; existing consistency proofs for maximum likelihood phylogenetics \cite{RoyChoudhury2015-ta} apply only to estimating model parameters when the ancestral sequences have been integrated out of the likelihood.
These proofs do not readily extend to include estimating ancestral states.
Moreover, examples of inconsistency arising from problems where the number of parameters increases with the amount of data \cite{Neyman1948-tt} indicate that joint inference of trees and ancestral states may not enjoy good statistical properties.
In this case those additional parameters come in the form of the states of ancestral sequences.
%EM I don't see this sentence in the most recent version of the paper via search. Can you point me to a section?
%das: changed give to given in case you were ctrl+f-ing, but it's here https://academic.oup.com/ve/article/4/1/vex042/4794731
%page 2 section 2.1 at the top of the second column first paragraph second sentence
Although \cite{Sagulenko2017-jo} explicitly warn that their approximation is for the case where ``branch lengths are short and only a minority of sites change on a given branch,'' their work motivates understanding when exactly this approximate technique breaks down, even when sequence data is ``perfect,'' i.e., is generated without sampling error according to the exact model used for inference.
